On mapping errors and seeding heuristics for DNA alignment

### 2019-04-10 MEM.tex
fixed typos sections 5 and 6

As mentioned above, after the terminator $\Downarrow_i$, we can append
segments with up to $\gamma+i-1$ symbols $\square$. The terminators of
those $\gamma+i$ possible segments are distributed according to the laws
of the arithmetic modulo $n+1$. For $i > 0$ for instance, if the next
segment has no symbol $\square$ then it is terminated by
$\Downarrow_{i-1}$. But the same goes if the segment has $n+1$ symbols
$\square$ (assuming $n+1 < \gamma$). Specifying the transfer matrix thus
involves a substantial amount of bookkeeping. One can check that the
expression of the transfer matrix $M_n(z)$ is
\begin{equation*}
\begin{blockarray}{cccccc}
   & \Dn{0} & \Dn{1} & \ldots & \Dn{n} & \nd \\
\begin{block}{c[ccccc]}
\Dn{0} & H_{0,0}(z) & H_{0,1}(z) & \ldots & H_{0,n}(z) & J_0(z) \\
\Dn{1} & H_{1,0}(z) & H_{1,1}(z) & \ldots & H_{1,n}(z) & J_1(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\Dn{n} & H_{n,0}(z) & H_{n,1}(z) & \ldots & H_{n,n}(z) & J_n(z) \\
\nd & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{eqnarray}
H_{i,j}(z) &=& pz \cdot (qz)^x \cdot \big( 1 + (qz)^{n+1} +
  \ldots + (qz)^{m(n+1)} \big), \\
\notag
  &\;& \text{where } x = i-j-1 \modulo{n+1},
  \text{ and } m = \left\lfloor
  \frac{\gamma+i-1-x}{n+1} \right\rfloor, \\
J_i(z) &=& 1 + qz + (qz)^2 + \ldots + (qz)^{\gamma+i-1}.
\end{eqnarray}

@@

$\tilde{M}_0(z)$ is too complex to compute a closed expression of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. It is actually easier to
proceed as in section~\ref{sec:example_skip} and to compute the powers of
$\tilde{M}_0(z)$ (using the arithmetic of truncated polynomials) up to a
given threshold value. Since each segment except the tail contains a
mismatch against at least one sequence, we define $\tilde{p}$ as the upper
bound on the probability of a mismatch $\tilde{p} = \max\{b,c,d\}$. The
updated formula (\ref{eq:bound}) now gives an upper bound of the
probability that a read of size $k$ contains $s$ or more mismatches as
\begin{equation*}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-\tilde{p})} -s\log
\frac{s}{k\tilde{p}} \right).
\end{equation*}

@@

The symbols $\downarrow_{/m}$ indicate that the nucleotide is a sequencing
error and $m \geq 0$ is the number of duplicates that \emph{match} the
nucleotide. Since a sequencing error is always a mismatch against the
target, the symbol $\downarrow_{/0}$ indicates that the ncleotide is a
mismatch against \emph{every} sequence. The symbols $\uparrow_{/i}$
indicate a change in masking status: the nucleotide is not masked but the
previous is. The index $i \geq 1$ is the number of nucleotides since the
last mismatch or since the beginning of the read. All the other
nucleotides are represented by the symbol $\square$ and the symbol $|$ is
appended to the end of the read as usual.

Note that in the symbols $\downarrow_{/m}$ and $\downarrow_{/i}$, the
numbers $m$ and $i$ have different meanings. In the symbol
$\downarrow_{/m}$, the index $m$ is a number of sequences ($0 \leq m \leq
N$ where $N$ is the number of duplicates); in the symbol $\uparrow_{/i}$,
the index $i$ is a number of nucleotides. Fig.~\ref{fig:sketch_extended}
shows the encoding of the read from Fig.~\ref{fig:exact_seed_example} in
the MEM alphabet.

@@

As before, the read is converted from a sequence of symbols to a sequence
of segments that consist of 0 or more symbols $\square$ followed by a
terminator. We then specify the weighted generating functions of those
segments and fill the tansfer matrix $\mathring{M}_N(z)$ of the reads that
do not contain an on-target MEM seed. We introduce the terms of the matrix
by increasing oder of complexity.

@@

The definition of $\omega_m$ is correct: if the nucleotide is an error,
the symbol is $\downarrow_{/m}$ for some $m$ between 0 and $N$. Each
duplicate is a match with probability $\mu/3$, so $m$ has a Binomial
distribution with parameters $(N, \mu/3)$.

@@

The symbol $\downarrow_{/m}$ signifies that the nucleotide has $m$ hard
masks and $N-m$ soft masks. If all the masks vanish before the first read
error, the next terminator will be a sybmol $\uparrow_{/j}$, otherwise it
will be the sybmol $|$ or a symbol $\downarrow_{/m}$. We separate the
cases based on the terminator of the segment.

@@

The definition of $\xi_j$ is correct: every nucleotide of the duplicate
differs from the target with probability $\mu$. In the absence of
sequencing error, this is also the probability that a nucleotide of the
duplicate differs from the read. Given that there is no error, the
probability that $j$ nucleotides in a row are identical to the read is
thus $(1-\mu)^j$ and the probability that at least one of them is
different is the complement $1-(1-\mu)^j$.

@@

Whenever a hard mask (here the first duplicate) covers the nucleotides as
shown by the grey squres in the left panel, there can be no on-target
seed. The positions marked with a question mark are irrelevant, they
cannot change the fact that there is no on-target MEM seed. If the hard
masks vanish, as in the right panel, then we need to look at the soft
masks. If a soft mask covers the whole segment as indicated by the grey
sqaures, then there can be no on-target seed. In all other cases there is
an on-target MEM seed.

@@

The definition of $\eta_j$ is correct: the probability that there are $j$
matches between the duplicate and the target is $(1-\mu)^j$. If there are
no sequencing errors, this is also the probability that there are $j$
matches between the duplicate and the read. The probability that the
duplicate matches the subsequent error is $\mu/3$, so the probability of
that there are $j+1$ matches including the sequencing error is
$(1-\mu)^j\mu/3$. Finally, the probability that there is a mismatch is the
complement $1-(1-\mu)^j\mu/3$.

Let us for now consider a segment of fixed size $j+1$. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability of condition $ii)$ is
\begin{equation*}
\xi_j^m \Big(1 - \eta_j^{N-m} \Big),
\end{equation*}
but we need to break up this term among all the possible terminators
$\downarrow_{/n}$ ($0 \leq n \leq N$) in order to fill the different
entries of the transfer matrix. For this, we split this term on the number
of soft masks that run until and including the terminator. From expression
(\ref{eq:eta}), the probability that there are $r \geq 1$ such soft masks
is
\begin{equation}
\label{eq:softmasks_r}
{N-m \choose r} (1 - \eta_j)^r \eta_j^{N-m-r}.
\end{equation}

@@

The case of the $N-m-r$ soft masks is more complicated because they can
vanish precisely on the terminator --- recall that in case $ii$) all the
hard masks are assumed to vanish before. If the soft mask failed within
the first $j$ nucleotides, then the $j+1$-th nucleotide can be anything
and it will match the terminator with probability $\mu/3$. But if the soft
mask survived the first $j$ nucleotides, then it \emph{must} fail on the
$j+1$-th and it cannot match the terminator. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability that a given soft mask
fails withtin the first $j$ nucleotides is $\xi_j/\eta_j$ --- this is the
conditional probability that it fails within the first $j$ nucleotides
given that it fails within the segment. Finally the probability that such
a soft mask matches the terminator is $\mu/3 \cdot \xi_j / \eta_j$.

@@

Summing the terms from case 3a and from case 3b, we find that the weighted
generating function of the $\downarrow_{/n}$ segments following
$\downarrow_{/m}$ is
\begin{equation}
\label{eq:A}
A_{m,n} =
\omega_n pz \sum_{j=0}^{\gamma-1} \Big(1 - \xi_j^N \Big) (qz)^j + \omega_n
pz \sum_{j=\gamma}^\infty \Big(1 - \xi_j^m \cdot
(1- \zeta_{j,m,n}) \Big) (qz)^j.
\end{equation}

@@

Collecting and arranging the results above, we can verify that the final
expression of the transfer matrix $\mathring{M}_N(z)$ is
\begin{equation*}
\begin{blockarray}{cccccccc}
   & \dn{0} & \ldots & \dn{N} & \up{1} & \ldots & \up{\gamma-1} & \nd \\
\begin{block}{c[ccccccc]}
\dn{0} & A_{0,0}(z) & \ldots & A_{0,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_0(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\dn{N} & A_{N,0}(z) & \ldots & A_{N,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_N(z) \\
\up{1} & D_{1,0}(z) & \ldots & D_{1,N}(z) & 0 & \ldots & 0 & E_1(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\up{\gamma-1} & D_{\gamma-1,0}(z) & \ldots & D_{\gamma-1,N}(z) & 0 &
  \ldots & 0 & E_{\gamma-1}(z) \\
\nd & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}
\end{equation*}
where
\begin{gather}
\tag{\ref{eq:A}}
A_{m,n} =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i \\
\tag{\ref{eq:B}}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i \\
\tag{\ref{eq:C}}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i \\
\tag{\ref{eq:D}}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i \\
\tag{\ref{eq:E}}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i
\end{gather}
and where
\begin{gather}
\tag{\ref{eq:omega}}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m \\
\tag{\ref{eq:xi}}
\xi_j = 1-(1-\mu)^j \\
\tag{\ref{eq:eta}}
\eta_j = 1-(1-\mu)^j\mu/3 \\
\tag{\ref{eq:zeta}}
\zeta_{j,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{rj} \psi_{j,m,n,r} \bigg/ {N \choose n} \\
\notag
\psi_{j,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_j^{n-r-q}.
\end{gather}

@@

The matrix $\mathring{M}_N(z)\cdot(I-\mathring{M}_N(z))^{-1}$ contains the
weighted generating functions of all the reads without on-target MEM
seeds. The term of interest, as usual, is the top right entry. Indeed,
every read can be prepended by $\downarrow_{/0}$ segments and only by
those, otherwise the read would start with fewer than $N$ soft masks.
Thus, reads without an on-target MEM seed are precisely the sequences of
segments that can be appended to the symbol $\downarrow_{/0}$ and that are
terimated by a tail.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_mem.pdf}
\caption{\textbf{Off target seeding probabilities (MEM seeds).}
The curves show the probability that seeding is off target for skip-9
seeds of size $\gamma=17$. Read size and erorr rate are the same as in
Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each line
shows a different number of duplicate sequences $N$ from 1 to 10 and the
x-axis shows the divergence rate $\mu$, defined as the probability that a
given duplicate differs from the target at any nucleotide.}
\label{fig:curves_mem}
\end{figure}

>>>>

As mentioned above, after the terminator $\Downarrow_i$, we can append
segments with up to $\gamma+i-1$ symbols $\square$. The terminators of
those $\gamma+i$ possible segments are distributed according to the laws
of the arithmetic modulo $n+1$. For $i > 0$ for instance, if the next
segment has no symbol $\square$ then it is terminated by
$\Downarrow_{i-1}$. But the same goes if the segment has $n+1$ symbols
$\square$ (assuming $n+1 < \gamma$). Specifying the transfer matrix thus
involves a substantial amount of bookkeeping. One can check that the
expression of the transfer matrix $M_n(z)$ is
\begin{equation*}
\begin{blockarray}{cccccc}
   & \Dn{0} & \Dn{1} & \ldots & \Dn{n} & \nd \\
\begin{block}{c[ccccc]}
\Dn{0} & H_{0,0}(z) & H_{0,1}(z) & \ldots & H_{0,n}(z) & J_0(z) \\
\Dn{1} & H_{1,0}(z) & H_{1,1}(z) & \ldots & H_{1,n}(z) & J_1(z) \\
\vdots & \vdots & \vdots & \ddots & \vdots & \vdots \\
\Dn{n} & H_{n,0}(z) & H_{n,1}(z) & \ldots & H_{n,n}(z) & J_n(z) \\
\nd & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray},
\end{equation*}
\begin{eqnarray}
H_{i,j}(z) &=& pz \cdot (qz)^x \cdot \big( 1 + (qz)^{n+1} +
  \ldots + (qz)^{m(n+1)} \big), \\
\notag
  &\;& \text{where } x = i-j-1 \modulo{n+1},
  \text{ and } m = \left\lfloor
  \frac{\gamma+i-1-x}{n+1} \right\rfloor, \\
J_i(z) &=& 1 + qz + (qz)^2 + \ldots + (qz)^{\gamma+i-1}.
\end{eqnarray}
where $\lfloor \ldots \rfloor$ denotes the ``floor'' function.

@@

$\tilde{M}_0(z)$ is too complex to allow computing a closed expression of
$\tilde{M}_0(z)\cdot(I-\tilde{M}_0(z))^{-1}$. It is actually easier to
proceed as in section~\ref{sec:example_skip} and to compute the powers of
$\tilde{M}_0(z)$ (using the arithmetic of truncated polynomials) up to a
given threshold value. Since each segment except the tail contains a
mismatch against at least one sequence, we define $\tilde{p}$ as the upper
bound on the probability of a mismatch $\tilde{p} = \max\{b,c,d\}$. The
updated formula (\ref{eq:bound}) now gives an upper bound of the
probability that a read of size $k$ contains $s$ or more mismatches as
\begin{equation*}
Pr(X \geq s) \leq \exp \left( (s-k)\log \frac{k-s}{k(1-\tilde{p})} -s\log
\frac{s}{k\tilde{p}} \right).
\end{equation*}

@@

The symbols $\downarrow_{/m}$ indicate that the nucleotide is a sequencing
error and $m \geq 0$ is the number of duplicates that \emph{match} the
nucleotide. Since a sequencing error is always a mismatch against the
target, the symbol $\downarrow_{/0}$ indicates that the nucleotide is a
mismatch against \emph{every} sequence. The symbols $\uparrow_{/i}$
indicate a change in masking status: the nucleotide is not masked but the
previous is. The index $i \geq 1$ is the number of nucleotides since the
last mismatch or since the beginning of the read. All the other
nucleotides are represented by the symbol $\square$ and the symbol $|$ is
appended to the end of the read as usual.

Note that in the symbols $\downarrow_{/m}$ and $\uparrow_{/i}$, the
numbers $m$ and $i$ have different meanings. In the symbol
$\downarrow_{/m}$, the index $m$ is a number of sequences ($0 \leq m \leq
N$ where $N$ is the number of duplicates); in the symbol $\uparrow_{/i}$,
the index $i$ is a number of nucleotides. Fig.~\ref{fig:sketch_extended}
shows the encoding of the read from Fig.~\ref{fig:exact_seed_example} in
the MEM alphabet.

@@

As before, the read is converted from a sequence of symbols to a sequence
of segments that consist of 0 or more symbols $\square$ followed by a
terminator. We then specify the weighted generating functions of those
segments and fill the transfer matrix $\mathring{M}_N(z)$ of the reads that
do not contain an on-target MEM seed. We introduce the terms of the matrix
by increasing order of complexity.

@@

The expression for $\omega_m$ is exact: if the nucleotide is an error,
the symbol is $\downarrow_{/m}$ for some $m$ between 0 and $N$. Each
duplicate is a match with probability $\mu/3$, so $m$ has a Binomial
distribution with parameters $(N, \mu/3)$.

@@

The symbol $\downarrow_{/m}$ signifies that the nucleotide has $m$ hard
masks and $N-m$ soft masks. If all the masks vanish before the first read
error, the next terminator will be a symbol $\uparrow_{/j}$, otherwise it
will be the symbol $|$ or a symbol $\downarrow_{/m}$. We separate the
cases based on the terminator of the segment.

@@

The expression for $\xi_j$ is exact: every nucleotide of the duplicate
differs from the target with probability $\mu$. In the absence of
sequencing errors, this is also the probability that a nucleotide of the
duplicate differs from the read. Given that there is no error, the
probability that $j$ nucleotides in a row are identical to the read is
thus $(1-\mu)^j$ and the probability that at least one of them is
different is the complement $1-(1-\mu)^j$.

@@

Whenever a hard mask (here the first duplicate) covers the nucleotides as
shown by the grey squres in the left panel, there can be no on-target
seed. The positions marked with a question mark are irrelevant, they
cannot change the fact that there is no on-target MEM seed. If the hard
masks vanish, as in the right panel, then we need to look at the soft
masks. If a soft mask covers the whole segment as indicated by the grey
squares, then there can be no on-target seed. In all other cases there is
an on-target MEM seed.

@@

The expression for $\eta_j$ is exact: the probability that there are $j$
matches between the duplicate and the target is $(1-\mu)^j$. If there are
no sequencing errors, this is also the probability that there are $j$
matches between the duplicate and the read. The probability that the
duplicate matches the subsequent error is $\mu/3$, so the probability of
that there are $j+1$ matches including the sequencing error is
$(1-\mu)^j\mu/3$. Finally, the probability that there is a mismatch is the
complement $1-(1-\mu)^j\mu/3$.

Let us for now consider a segment of fixed size $j+1$. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability of condition $ii)$ is
\begin{equation*}
\xi_j^m \Big(1 - \eta_j^{N-m} \Big),
\end{equation*}
but we need to break up this term among all the possible terminators
$\downarrow_{/n}$ ($0 \leq n \leq N$) in order to fill the different
entries of the transfer matrix. For this, we split this term in the number
of soft masks that run until and including the terminator. From expression
(\ref{eq:eta}), the probability that there are $r \geq 1$ such soft masks
is
\begin{equation}
\label{eq:softmasks_r}
{N-m \choose r} (1 - \eta_j)^r \eta_j^{N-m-r}.
\end{equation}

@@

The case of the $N-m-r$ soft masks is more complicated because they can
vanish precisely on the terminator --- recall that in case $ii$) all the
hard masks are assumed to vanish before. If the soft mask failed within
the first $j$ nucleotides, then the $j+1$-th nucleotide can be anything
and it will match the terminator with probability $\mu/3$. But if the soft
mask survived the first $j$ nucleotides, then it \emph{must} fail on the
$j+1$-th and it cannot match the terminator. From expressions
(\ref{eq:xi}) and (\ref{eq:eta}), the probability that a given soft mask
fails within the first $j$ nucleotides is $\xi_j/\eta_j$ --- this is the
conditional probability that it fails within the first $j$ nucleotides
given that it fails within the segment. Finally the probability that such
a soft mask matches the terminator is $\mu/3 \cdot \xi_j / \eta_j$.

@@

Summing the terms from case 3a and from case 3b, we find that the weighted
generating function of the $\downarrow_{/n}$ segments following
$\downarrow_{/m}$ is
\begin{equation}
\label{eq:A}
A_{m,n}(z) =
\omega_n pz \sum_{j=0}^{\gamma-1} \Big(1 - \xi_j^N \Big) (qz)^j + \omega_n
pz \sum_{j=\gamma}^\infty \Big(1 - \xi_j^m \cdot
(1- \zeta_{j,m,n}) \Big) (qz)^j.
\end{equation}

@@

Collecting and arranging the results above, we can verify that the final
expression of the transfer matrix $\mathring{M}_N(z)$ is
\begin{equation*}
\begin{blockarray}{cccccccc}
   & \dn{0} & \ldots & \dn{N} & \up{1} & \ldots & \up{\gamma-1} & \nd \\
\begin{block}{c[ccccccc]}
\dn{0} & A_{0,0}(z) & \ldots & A_{0,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_0(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\dn{N} & A_{N,0}(z) & \ldots & A_{N,N}(z) & B_1(z) & \ldots &
    B_{\gamma-1}(z) & C_N(z) \\
\up{1} & D_{1,0}(z) & \ldots & D_{1,N}(z) & 0 & \ldots & 0 & E_1(z) \\
\vdots & \vdots & \ddots & \vdots & \vdots & \ddots &
    \vdots & \vdots \\
\up{\gamma-1} & D_{\gamma-1,0}(z) & \ldots & D_{\gamma-1,N}(z) & 0 &
  \ldots & 0 & E_{\gamma-1}(z) \\
\nd & 0 & \ldots & 0 & 0 & \ldots & 0 & 0 \\
\end{block}
\end{blockarray}
\end{equation*}
where
\begin{gather}
\tag{\ref{eq:A}}
A_{m,n}(z) =
\omega_n pz \sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i + \omega_n
pz \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \cdot
(1- \zeta_{i,m,n}) \Big) (qz)^i \\
\tag{\ref{eq:B}}
B_i(z) = \Big( \xi_i^N-\xi_{i-1}^N \Big) (qz)^i \\
\tag{\ref{eq:C}}
C_m(z) =
\sum_{i=0}^{\gamma-1} \Big(1 - \xi_i^N \Big) (qz)^i +
  \sum_{i=\gamma}^\infty \Big(1 - \xi_i^m \Big) (qz)^i \\
\tag{\ref{eq:D}}
D_{j,m}(z) = \omega_m pz \sum_{i=0}^{\gamma-j-1} (qz)^i \\
\tag{\ref{eq:E}}
E_j(z) = \sum_{i=0}^{\gamma-j-1} (qz)^i
\end{gather}
and where
\begin{gather}
\tag{\ref{eq:omega}}
\omega_m = {N \choose m} \big(1 - \mu/3\big)^{N-m} \big(\mu/3\big)^m \\
\tag{\ref{eq:xi}}
\xi_j = 1-(1-\mu)^j \\
\tag{\ref{eq:eta}}
\eta_j = 1-(1-\mu)^j\mu/3 \\
\tag{\ref{eq:zeta}}
\zeta_{j,m,n} = \sum_{r\geq1} {N-m \choose r}
(1-\mu)^{rj} \psi_{j,m,n,r} \bigg/ {N \choose n} \\
\notag
\psi_{j,m,n,r} = \sum_{q \geq 0}{m \choose q}{N-m-r \choose n-r-q}
\xi_j^{n-r-q}.
\end{gather}

@@

The matrix $\mathring{M}_N(z)\cdot(I-\mathring{M}_N(z))^{-1}$ contains the
weighted generating functions of all the reads without on-target MEM
seeds. The term of interest, as usual, is the top right entry. Indeed,
every read can be prepended by $\downarrow_{/0}$ segments and only by
those, otherwise the read would start with fewer than $N$ soft masks.
Thus, reads without an on-target MEM seed are precisely the sequences of
segments that can be appended to the symbol $\downarrow_{/0}$ and that are
terminated by a tail.

@@

\begin{figure}[h]
\centering
\includegraphics[scale=0.65]{curves_mem.pdf}
\caption{\textbf{Off target seeding probabilities (MEM seeds).}
The curves show the probability that seeding is off target for skip-9
seeds of size $\gamma=17$. Read size and error rate are the same as in
Fig.~\ref{fig:curves_exact}, \textit{i.e.} $k=50$ and $p=0.01$. Each line
shows a different number of duplicate sequences $N$ from 1 to 10 and the
x-axis shows the divergence rate $\mu$, defined as the probability that a
given duplicate differs from the target at any nucleotide.}
\label{fig:curves_mem}
\end{figure}
